# R Startup

**Learning objectives:**

-   Set environment variables in `.Renviron`.
-   Run R code at R startup with `.Rprofile`.
-   Run R without startup files.

## .Renviron {-}

My `.Renviron` looks something like this:

```
SLACK_SKIPLOAD=TRUE
R_DEFAULT_INTERNET_TIMEOUT=1000
NOT_CRAN=true
RTOOT_DEFAULT_TOKEN=<TOKEN_REDACTED>;user;fosstodon.org
LINKEDIN_TOKEN=<TOKEN_REDACTED>
OPENAI_API_KEY=<TOKEN_REDACTED>
```

Project-specific versions can be useful for [GitHub Actions & shinyapps.io deployments](https://github.com/r4ds/mentordash/blob/main/.github/workflows/deploy_shinyapps.yml).

## .Rprofile {-}

My `.Rprofile` got crazy enough that I made [`{andthis}`](https://github.com/jonthegeek/andthis) for my customizations:

```{r rprofile, eval = FALSE}
# Set this to FALSE (or comment out) to update packages like usethis.
options(jon_load_stuff = TRUE)

# Set package options. Make sure these don't impact how analyses work!
options(
  usethis.full_name = "Jon Harmon",
  usethis.description = list(
    `Authors@R` = 'person("Jon", "Harmon",
                          email = "jonthegeek@gmail.com",
                          role = c("aut", "cre"),
                          comment = c(ORCID = "0000-0003-4781-4346"))',
    License = "MIT + file LICENSE"
  ),
  usethis.destdir = "C:/Users/jonth/Dropbox (Personal)/R"
)
options(tidymodels.dark = TRUE)
options(styler.cache_root = "styler-perm")

# Set options that make coding safer.
options(
  warnPartialMatchArgs = TRUE,
  warnPartialMatchDollar = TRUE,
  warnPartialMatchAttr = TRUE
)

# Set up my dev environment (load packages, etc).
if (interactive() && getOption("jon_load_stuff", FALSE)) {
  suppressMessages(require(devtools))
  suppressMessages(require(reprex))
  suppressMessages(require(andthis))

  # Show branch + time in console prompt.
  prompt::set_prompt(
    function(...) {
      paste0(
        "[",
        prompt::git_branch(),
        " ",
        format(Sys.time(), "%T"),
        "] > "
      )
    }
  )
}
```

## R without Startups {-}

If running R from command line:

-   `--vanilla`
    -   `.Rprofile` `r emoji::emoji("x")`
    -   `.Renviron`  `r emoji::emoji("x")`
-   `--no-init-file` (general) or `--no-site-file` (site?)
    -   `.Rprofile` `r emoji::emoji("x")`
    -   `.Renviron`  `r emoji::emoji("check_mark")`
-   `--no-environ`
    -   `.Rprofile` `r emoji::emoji("check_mark")`
    -   `.Renviron`  `r emoji::emoji("x")`

Can also disable `.Rprofile` in project options in Rstudio.

## Wait what's a "site" `Rprofile`? {-}

Mine is here: `C:\Program Files\R\R-4.2.2\etc\Rprofile.site`

It's set up with defaults and notes (I haven't edited):

```{r Rprofile.site, eval = FALSE}
# Things you might want to change

# options(papersize="a4")
# options(editor="notepad")
# options(pager="internal")

# set the default help type
# options(help_type="text")
  options(help_type="html")

# set a site library
# .Library.site <- file.path(chartr("\\", "/", R.home()), "site-library")

# set a CRAN mirror
# local({r <- getOption("repos")
#       r["CRAN"] <- "http://my.local.cran"
#       options(repos=r)})

# Give a fortune cookie, but only to interactive sessions
# (This would need the fortunes package to be installed.)
#  if (interactive()) 
#    fortunes::fortune()
```

## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/8vROEoMhZV8")`

<details>
<summary> Meeting chat log </summary>

```
00:13:43	Jon Harmon (jonthegeek):	usethis::edit_r_environ()
00:24:09	Jon Harmon (jonthegeek):	usethis::edit_r_profile()
00:25:31	Jon Harmon (jonthegeek):	use_conflicted()
use_reprex()
use_usethis()
use_devtools()
use_partial_warnings()
00:26:32	Jon Harmon (jonthegeek):	>
00:27:03	Jon Harmon (jonthegeek):	[main 11:42:50] > edit_r_profile()
• Modify 'C:/Users/jonth/Documents/.Rprofile'
• Restart R for changes to take effect
[main 13:25:05] >
00:32:20	Jon Harmon (jonthegeek):	[main 13:25:05] > getOption("repos")
                       CRAN 
"https://cran.rstudio.com/" 
attr(,"RStudio")
[1] TRUE
[main 13:30:21] >

```
</details>
